### 2023-07-18 paper/appendix/dataset.tex

\section{Appendix -- Dataset Details}
The following details in this appendix are for clarification purposes.
All examples in this appendix are created via PySpark using dataframes.

\subsection{Using PySpark to Load Data}
All example code was created using Jupyter Notebooks creating a PySpark session. The deails of setting up such session are below in the code example. All subsequent details are derived from the \texttt{sc}\footnote{sc = Spark Session}.
\begin{minted}{python}
# Setup Spark Session
# Spark Session Name: Qualyfing_Exam
from pyspark.sql import SparkSession

# Get Imports Needed
from pyspark.sql.functions import col, udf

# Get Datatypes needed for DataFrame manipulation
from pyspark.sql.types import IntegerType, StringType

# Setup Spark Session
sc = SparkSession \
        .builder \
        .master("local[*]") \
        .appName("Qualyfing_Exam") \
        .getOrCreate()
\end{minted}

\subsection{Details on English Dataset}
The English dataset based on tweets\footnote{Tweets from Twitter} are saved as a \texttt{csv}\footnote{csv = Comma Separated Values} file. The following are some simply details on the laod and processing of such file.
\begin{minted}{python}
# Read csv file to Saprk dataframe
df_english = sc.read.csv("data/english.csv", header=True, inferSchema=True)
# Group by a column and count the occurrences in df_english
count_english = df_english.groupBy('sentiment').count()

# Show the result for df_english
count_english.show()
+----------+-----+
| sentiment|count|
+----------+-----+
|   boredom|  179|
|      love| 3842|
|    relief| 1526|
|       fun| 1776|
|      hate| 1323|
|   neutral| 8638|
|     anger|  110|
| happiness| 5209|
|  surprise| 2187|
|   sadness| 5165|
|     worry| 8459|
|enthusiasm|  759|
|     empty|  827|
+----------+-----+

\end{minted}

\subsection{Details on German Dataset}
The German dataset was downloaded from the ETH's\footnote{ETH (German: Eidgenössische Technische Hochschule) or Swiss Federal Institute of Technology} servers via "link" as a \texttt{JSON} file.
\begin{minted}{python}
# Read JSON file to Saprk dataframe
df_german = sc.read.json("data/german.json")

# Group by a column and count the occurrences in df_german
count_german = df_german.groupBy('article_emotion').count()

# Show the result for df_german
count_german.show(truncate=0)
+----------------------------------+-----+
|article_emotion                   |count|
+----------------------------------+-----+
|[Ärger, Antizipation, Traurigkeit]|1    |
|[Überraschung, Vertrauen]         |14   |
|[Überraschung, Freude]            |15   |
|[Ärger, Vertrauen]                |7    |
|[Ekel, Traurigkeit]               |4    |
|[Keine, Antizipation]             |5    |
|[Vertrauen]                       |130  |
|[Freude]                          |58   |
|[Ärger, Antizipation]             |31   |
|[Ärger]                           |114  |
|[Überraschung]                    |154  |
|[Antizipation, Traurigkeit]       |15   |
|[Antizipation, Freude, Vertrauen] |2    |
|[Vertrauen, Angst]                |5    |
|[Traurigkeit]                     |101  |
|[Antizipation]                    |413  |
|[Traurigkeit, Vertrauen]          |7    |
|[Antizipation, Vertrauen]         |104  |
|[Ärger, Traurigkeit, Angst]       |1    |
|[Unklar]                          |314  |
+----------------------------------+-----+

\end{minted}

\clearpage