### 2023-07-18 paper/appendix/code.tex
Paper Updates

\section{Appendix -- Application \& Code Details}
Hello World - Application \& Code Details
\subsection{Application}
\label{appendix:code_application}
Hello World - Application
\clearpage

\subsection{Code Details}
Hello World - Code Details

\subsection{RESTful API Server}
\label{appendix:code-api-server}
The RESTful\footnote{RESTful API is an interface that two computer systems use to exchange information securely over the internet.} API Server will be a simple Python Flask server listening on port \texttt{8080} on the local machine.

\clearpage