### 2023-07-31 paper/main.tex

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[toc,page]{appendix}
\usepackage{microtype}
\usepackage{fullpage}
\usepackage{multirow}

\usepackage{minted}
\setminted[]{linenos, numbersep=5pt, frame=lines}

\usepackage{parskip}
\setlength{\parindent}{0em}
\setlength{\parskip}{1em}

\usepackage{array}
\newcolumntype{L}{>{\centering\arraybackslash}m{1.2in}}

\usepackage[affil-it]{authblk}
\usepackage{amsfonts, amssymb, amsmath}

\usepackage{wrapfig}
\usepackage{graphicx}
\graphicspath{ {./images/} }

\usepackage{float}

\usepackage[backend=biber]{biblatex}
\addbibresource{sources.bib}
% Use the followign Command to validate *.bib file: biber --tool --validate-datamodel qualifying_exam_paper.bib

\title{Improving Emotion Detection through Real-Time Translation of Text to ML Models Trained in Different Languages}

\author{Richard Hoehn%
	\thanks{Electronic address: \texttt{rhoehn@mtmail.mtsu.edu}; corresponding author}}
\affil{Middle Tennessee State University}

\author{Dr. Jaishree Ranganathan%
	\thanks{Electronic address: \texttt{jaishree.ranganathan@mtsu.edu}}}
\affil{Middle Tennessee State University}

\begin{document}

\maketitle

\begin{abstract}
This Qualifying Exam\footnote{PhD students of MTSU Computational and Data  Science program are required to complete a Qualifying Exam in their first year in the program.} research paper investigates the potential of improving Emotion Detection (ED) through real-time translation of text to multiple Machine Learning (ML) models trained in different languages (English and German). The research tries to address the challenges posed by the lack of comprehensive labeled datasets and language fragmentation in ED analysis.

By extending an original language dataset with a translated dataset, in the case of this paper English and German. And utilizing parallel processing with PySpark, this research project aims to achieve higher prediction rates in ED. The results presented both orally to the MTSU's Computational Data \& Science Committee and within this paper demonstrate the effectiveness of such an approach and how the real-time translation technique can enhance emotion detection capabilities.
\end{abstract}
\clearpage

\tableofcontents
\clearpage

\section{Introduction}

Emotion detection has gained significant interest in recent years as a means to understand human behavior and improve communication effectiveness. By analyzing text or speech, these detection processes can accurately identify and classify emotions expressed by individuals or groups, often times in real-time.

By employing a variety of ED processes, including natural language processing and machine learning, applications can detect emotions with high precision. These ED systems find applications in diverse fields, such as sentiment analysis, customer feedback analysis, voting and news stances, and mental health monitoring, which all aim at contributing to a deeper understanding of human emotions and their impact on various aspects of life.

\subsection{Human Emotion as a Theoretical Framework}

The concept and number of emotions can vary depending on the theoretical framework or model being considered. One well-known model is the Plutchik's Wheel of Emotions Figure \ref{fig:plutchik-wheel}, which proposes eight (8) primary emotions and their contrasting pairs. In this paper we will be focusing on the basic eight emotions and not on their sub-emotions.

Below are the eight primary emotions according to the Plutchik's Wheel of Emotions \cite{Tromp}, a detailed reference of emotions used in this research can be found in Table \ref{table:dataset_labels} on page~\pageref{table:dataset_labels}.

\begin{itemize}
\item Happiness: A feeling of Joy (Fun), contentment, or delight.
\item Sadness: A state of unhappiness, sorrow, or grief.
\item Anger: An intense emotional response associated with feelings of displeasure, frustration, or hostility.
\item Worry: An emotional response triggered by perceived threats or danger, often associated with anxiety or terror.
\item Surprise: A sudden and unexpected reaction to something unusual or startling.
\item Disgust and/or Hate: A strong aversion or revulsion towards something unpleasant, offensive, or repulsive.
\item Trust and/or Love: A sense of reliance, confidence, or faith in someone or something.
\item Enthusiasm: A feeling of excited expectation or eagerness towards future events.
\end{itemize}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.4\textwidth]{Plutchik-Wheel}
    \caption{Plutchik Wheel of Emotions}
    \label{fig:plutchik-wheel}
\end{figure}

It's important to note that emotions are complex and multifaceted, and different models may propose alternative categorizations or variations in the number of primary emotions. However for the scope of this research project the most common eight (8) emotions are used as described above.

By aligning the emotional labels across the English \cite{english-dataset-twitter} and German \cite{german-dataset-cheese} datasets being used in this research all eight listed above were used. Details on the linkage of the two datasets is described in Section \ref{sec:data-procurement-process} with an emphasis on the Table \ref{table:dataset_labels}'s cross reference on page~\pageref{table:dataset_labels}.

\subsection{Significance of Emotion Detection (ED) in Machine Learning}
Emotion Detection (ED) analysis in Machine Learning (ML) models has gained significant attention due to its potential applications in various fields, including customer sentiment analysis, stance detection in regard to a specific targets such as news and voting\cite{mascarell-etal-2021-stance}, mental health monitoring\cite{Colonnello}, and human-computer interactions (chatbots\cite{chatbot-cognitive-awareness}) to just name a few.

By accurately identifying and understanding emotions from text data, ML applications can assist in improving user experiences, decision-making processes, and overall human-machine interactions in a positive manner\cite{Colonnello, mascarell-etal-2021-stance}. As an example, the primary goals of chatbots are entertainment, social contact, and novelty interaction, with a strong emphasis on productivity, scale-ability and cost reductions. In chatbots, there has been a immense requirement to simulate human-like characteristics and behavior during human-machine interactions\cite{emotion-detection-literature-review}. Per Kusal et al. \cite{chatbot-cognitive-awareness} these chatbots also improve customer engagement by offering friendliness, comfortness, flexibility, and efficient assistance. Chatbots should provide customers with more engaging responses, directly addressing their issues. The goal of deploying ED applications is for users perceive chatbots as companions rather than simple assistants. Per Adikari et al. the majority of user requests are emotional rather than informative. Chatbots gained the capacity to respond emotionally to customers due to machine learning and sentiment analysis evolution\cite{chatbot-cognitive-awareness}.

\begin{quote}
    I usually know almost exactly how I feel. The problem is, I just can’t tell anyone.
    \flushright -- Meg Cabot
\end{quote}

Due to the before mentioned wide scope of uses for ED this field of research and the work to advance it provides significance and importance in enabling machines to comprehend and respond appropriately to human interactions. This in turn contributes to the advancement of artificial intelligence, natural language processing and in many cases the utilization of text-to-speech applications most all of us in today's society use and rely on \cite{ai-framework-detection-emotions}.

\subsection{ED challenges: Lack of Comprehensive Labeled Datasets and Language Fragmentation}
ED is an immensely growing field, however unlike Sentiment Analysis (SA) the availability of large datasets for training purposes of ML models is much smaller \cite{ACLU-ED-Data, ai-framework-detection-emotions}. The primary reason of the lack of data stems from the emotion detection data requiring primarily supervised learning dataset, which require time and effort to procure, and mostly created and labeled by humans. To compound on the issue of the lack of data; the many datasets that are available are in multiple languages. A significant amount of time took place to find the German labeled

\subsection{Motivation for the research project focusing on extending ED datasets through translation and evaluating the impact on prediction rates}

This paper is significant because an entire industry of automated emotion reading (Text, Image, Video, and Sound) technologies is quickly emerging \cite{ACLU-ED-Data, ACLU-THE-DAWN-OF-ROBOT-SURVEILLANCE}. The market for emotion recognition software is forecast to reach at least \$3.8 billion by 2025 \cite{ACLU-ED-Data, ACLU-THE-DAWN-OF-ROBOT-SURVEILLANCE}. And since ED is already being incorporated into products for purposes such as marketing, robotics, driver safety, customer service and a multiple of others, it is fitting to work on research to solve issues with prediction rates and lack of teaching / learning data.

With the before mentioned issues regarding the quantity (lack of) and fragmentation of the ED datasets this research was motivated by finding a way to combine different language datasets into a single set for training processes to improve the predictability of ED. The hope is that by use of translation that the predictability of ED can be improved by use of real-time language conversion and processing.

\subsection{Objectives \& Scope of the Research}

The research project aims at answering the following research questions:

\begin{itemize}
\item Can by translating English data to German and extending an original German dataset increase the predictability of the ED model?
\item Can by translating German data to English and extending an original English dataset increase the predictability of the ED model?
\item Can by translating in real-time an input to multiple languages improve the predictability based on the combined output of two model's results.
\end{itemize}

\section{Literature Review}
To ensure comprehensive coverage of the latest developments and research in ED for this research paper, consideration and review was only given to incorporating those research papers and online news articles that were published after the year 2018, thereby providing an up-to-date perspective on the subject matter.

\subsection{Emotion Detection Importance and Diverse Populations}
Emotion Detection (ED) holds significant importance in various domains, including psychology, social sciences, and human-computer interaction. It enables the analysis of human emotions at scale, providing valuable insights into individual and collective emotional states.

Furthermore, emotions can differ across age groups, genders, cultures, and languages. Including data from diverse populations helps in developing inclusive and culturally sensitive ED models. It ensures that the models are not biased towards specific groups and can accurately detect emotions in a wide range of individuals.

\subsection{Lack of Labeled Datasets and the Fragmentation caused by Different Languages}

The lack of diverse and large-scale data hinders the development and training of accurate ED models, limiting their effectiveness and general ability across different contexts and populations.

In recent research, efforts have been made to address these limitations by exploring new techniques for emotion detection. For instance, a study by Kapoor\cite{KAPOOR2023120882} et al. proposes analyzing the entire spectrum of the language source to predict emotional changes. This provided some help, but quickly established itself as time consuming and need of expertise in the subject matters. Overcoming the scarcity of labeled datasets and embracing innovative approaches like the before mentioned study by Kapoor\cite{KAPOOR2023120882} et al. and their analysis can contribute to the advancement of ED research. Such advancements hold the potential to enhance the accuracy and applicability of emotion detection models, enabling a deeper understanding of human emotions and more effective responses in various domains.

The paper by Mascarell\cite{mascarell-etal-2021-stance} et al. states that unlike images language operate very differently, therefore the lack of labeled datasets poses a significant challenge in text-based Emotion Detection (ED) analysis as it limits the availability of reliable training data for ML models. This scarcity hampers the model's ability to learn and generalize emotions effectively.

Additionally, Kusal et al. further stipulates \cite{kusal} that the fragmentation caused by different languages further exacerbates the issue, as it reduces the size and diversity of data available for training, resulting in limited cross-lingual generalization and potentially biased models. Overcoming these challenges is crucial for advancing ED research and improving the accuracy and applicability of emotion detection models across various languages and cultures.

\section{Methodology}

\subsection{Data procurement process of ED datasets in English and German}
\label{sec:data-procurement-process}
The data procurement was relatively straight forward and in essence entailed using Google search terms for the Emotion Detection in English and German. Some of the main German data was obtained from the dataset built by ETH's Emotion and Stance Detection for German Text \cite{mascarell-etal-2021-stance} which also hosts a good website with their findings \footnote{https://mtc.ethz.ch/research/natural-language-processing/emotion-stance.html}. The English dataset was downloaded from Kaggle \footnote{https://www.kaggle.com/datasets/pashupatigupta/emotion-detection-from-text} based on Tweets collected in 2021. The English dataset contains over 38,000 tweets and their corresponding emotion labeled. Similarly the German contains over 12,000 sentences and their respective emotions associated.

An important part of the process is to make sure we match the English emotions to those of the German language emotions. Since the tow languages are dissimilar we opted to match exact emotions to each other that can be seen in Table \ref{table:dataset_labels} below.

\begin{table}[h!]
\centering
\begin{tabular}{ | c c | c c | c | }
    \hline
    \multicolumn{5}{|c|}{Emotion Datasets Labels} \\

    \hline
    \multicolumn{2}{|c|}{English} & \multicolumn{2}{|c|}{German} & \multirow{2}{*}{Used} \\
    Name & Count & Name & Count \\
    \hline
    Boredom    &  179 & ---           & --- & NO  \\
    Love       & 3842 & Vertrauen     & 316 & YES \\
    Relief     & 1526 & ---           & --- & NO  \\ 
    Fun        & 1776 & ---           & --- & NO  \\
    Hate       & 1323 & Ekel          &  29 & YES \\
    Neutral    & 8638 & Unklar        & 314 & YES \\
    Anger      &  110 & Ärger         & 226 & YES \\
    Happiness  & 5209 & Freude        & 140 & YES \\
    Surprise   & 2187 & Überraschung  & 369 & YES \\
    Sadness    & 5165 & Traurigkeit   & 184 & YES \\
    Worry      & 8459 & Angst         & 154 & YES \\
    Enthusiasm &  759 & Antizipation  & 774 & YES \\
    Empty      &  827 & ---           & --- & NO  \\
    \hline
\end{tabular}
\caption{Dataset of English \& German and their respective label counts.}
\label{table:dataset_labels}
\end{table}

Details on how we formatted the the two (English and German) datasets to each other can be viewed in the Appendix \ref{appendix:dataset_english} and \ref{appendix:dataset_german}

The procurement of two ED datasets is needed. The first will be in English and the second in German. Each dataset will be translated into the other’s language for extending purposes. As an example Figure \ref{fig:dataset-extension} depicts the original English dataset will be extended by the translated German dataset.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.35\textwidth]{Dataset-Extension}
    \caption{Example of English dataset extension by translating from German.}
    \label{fig:dataset-extension}
\end{figure}

The reason for selecting German was based on the author\footnote{Richard Hoehn} being fluent in English and German and hence being able to easily traverse the two languages while working on the code for dataset translations and formatting.

Once the two datasets have been translated, we now extend the original language dataset with the translated one.By this approach we are in essence extending the original datasets to create a larger one with the foreign language data. With both extended datasets we can now train our ML models for English and German.

\subsection{Development of a translation application for converting different languages into English and vice versa}
Translation of the emotion datasets was performed with Google Translator based on the Medium article\cite{Nidhaloff_how_to_translate_text_with_python} in which the author describes option for simple API translations. As an example the following is a brief introduction to Google Translator's API:

\begin{quote}
\begin{minted}[fontsize=\small]{python}
# Using Deep Translator to leveage Google Translate
# Link: https://cloud.google.com/translate/docs/reference/libraries/v2/python
from deep_translator import GoogleTranslator

sentence = 'Chocolate milk is so much better through a straw.'

translated = GoogleTranslator(source='auto', target='de').translate(sentence)
print(translated) # Schokoladenmilch schmeckt durch einen Strohhalm viel besser.

translated_back = GoogleTranslator(source='auto', target='en').translate(translated)
print(translated_back) # Chocolate milk tastes much better through a straw.
\end{minted}
\end{quote}

Based on the above example the CSV and JSON datasets were individually loaded, parsed, and translated to their respective counter-part languages. Details of the application can be found in the Appendix \ref{appendix:dataset_english} for English and Appendix \ref{appendix:dataset_german} for the German. We decided to only use 1,500 row for each language so since the German (DE\footnote{DE = German}) dataset was not much larger than that. This meant that the English (EN\footnote{EN = Enlgish}) datasets was randomly paired down to 1,500 to match that of the German.

Each translated file now contains four (4) columns that are labeled as in Tables \ref{table:pd_de_translated} and \ref{table:pd_en_translated} with the exact same column naming.

\begin{table}[h!]
\centering
\begin{tabular}{ | L | L | L | L | }
    \hline
    \multicolumn{4}{|c|}{File: \texttt{pd\textunderscore de\textunderscore translated.csv}} \\
    \hline
    sentence\textunderscore de &
    emotion\textunderscore de & 
    sentence\textunderscore en & 
    emotion\textunderscore en \\
    \hline
    Original Sentence (feature) &
    Original Emotion (label) &
    Sentence Translated to English & 
    Emotion set by Cross-Reference from German \\

    \hline
\end{tabular}
\caption{German CSV file structure after translation to English}
\label{table:pd_de_translated}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{ | L | L | L | L | }
    \hline
    \multicolumn{4}{|c|}{File: \texttt{pd\textunderscore en\textunderscore translated.csv}} \\
    \hline
    sentence\textunderscore en &
    emotion\textunderscore en & 
    sentence\textunderscore de & 
    emotion\textunderscore de \\
    \hline
    Original Sentence (feature) &
    Original Emotion (label) &
    Sentence Translated to German & 
    Emotion set by Cross-Reference from English \\

    \hline
\end{tabular}
\caption{English CSV File structure after translation to German}
\label{table:pd_en_translated}
\end{table}

\subsection{Training Multiple ML models using the Extended Datasets with PySpark}

 Both sentences are then processed in real-time on both the English and German models for the Emotion Detection and will be compared to each other's output. The average of the prediction will be used for final prediction of the supplied sentence and the emotion that was detected.

 \begin{table}[h!]
\centering
\begin{tabular}{ | L | L | L | L | }
    \hline
    \multicolumn{4}{|c|}{ML Models } \\
    \hline
    Model & 
    Data &
    Rows for Training & 
    Rows for Testing  \\
    \hline
    A &
    English Original &
    1200 & 
    300 (20\%) \\
    \hline
    B &
    English Extended By German &
    2700 & 
    300 same as Model "A" \\
    \hline
    C &
    German Original &
    1200 & 
    300 (20\%) \\
    \hline
    D &
    German Extended By English &
    2700 & 
    300 same as Model "C" \\

    \hline
\end{tabular}
\caption{ML Models and their respective settings}
\label{table:ml_models}
\end{table}

In order for such to be processed in real-time, the Jupyter Notebook will start a PySpark session that is accessed via the local server running that accepts RESTful API calls with details of the server in Appendix \ref{appendix:code-api-server}.

\subsection{Evaluation Metrics and methodologies for measuring the performance of the ML models}
The evaluation of ML models trained on text datasets is a critical step in assessing their performance and reliability. Our research project is no different and the following assumptions are used to determine the effectiveness of the methodology of extending our datasets:
\begin{itemize}
\item \textbf{Accuracy:} Accuracy measures the overall correctness of the emotion predictions made by the model. It is calculated as the ratio of correctly classified instances to the total number of instances in the dataset. However, accuracy may not provide a complete picture, especially if the dataset is imbalanced.
\item \textbf{Precision, Recall, and F1 Score:} Precision represents the proportion of correctly predicted emotions among all instances predicted as a specific emotion. Recall measures the proportion of correctly predicted emotions among all instances that truly belong to a specific emotion. F1 score combines precision and recall into a single metric, providing a balanced evaluation.
\item \textbf{Cross-Validation:} Cross-validation techniques, such as k-fold or stratified cross-validation, divide the dataset into multiple subsets for training and testing. This helps evaluate the model's performance across different data samples, mitigating over-fitting and providing a more robust estimation of its generalization ability.
\item \textbf{Error Analysis:} Performing error analysis on the model's predictions helps identify specific patterns or instances where the model might struggle or make frequent mistakes. This analysis can guide improvements and provide insights for future model development.
\end{itemize}

When evaluating such models, several factors come into play. First and foremost, the choice of appropriate evaluation metrics is crucial, depending on the specific task at hand, such as text classification, sentiment analysis, or named entity recognition. Metrics like accuracy, precision, recall, F1 score, or area under the receiver operating characteristic curve (AUC-ROC) are commonly used to measure the models' performance.

\subsubsection{Written \& Oral Presentation of Research Finding}
In conclusion of the project a research paper and oral presentation is required to pass MTSU Computation Science PhD Qualifying Exam. Such paper is being presented with this paper and the oral presentation will be in Microsoft Power Point format.

The listing the datasets used, code for the application, and details of the research project process in this research paper is openly available and hosted on a GitHub repository, allowing for easy access, reproducible, and collaborative engagement with the research findings. The repository is open and available under: \texttt{https://github.com/richardhoehn/mtsu.coms.qualifying-exam}\cite{Hoehn_Improving_Emotion_Detection_2023}

\section{Results and Analysis}

\subsection{Presentation of the prediction results obtained from the extended datasets}

\ldots

\subsection{Comparison of Prediction Result of the Original and Extended Language (translated) Datasets}
\ldots

\subsection{Analysis of the impact of Extending Datasets by Translation and Performance of the Combination of ML models}
\ldots

\subsection{Discussion of the findings and their implications for improving Emotion Detection}
\ldots

\section{Demonstration and Application}

\subsection{Description of the developed translation application for real-time text translation}
Details of the entire application can be found on the github repository \cite{Hoehn_Improving_Emotion_Detection_2023}.

\subsection{Showcase of the application's functionality and its integration with the Emotion Detection process}
\ldots

\subsection{Demonstration of the real-time processing of translated text and emotion detection using the ML models}
\ldots

\subsection{GIT Repository Details}
The listing the datasets used, code for the application, and details of the research project process in this research paper is openly available and hosted on a GitHub repository, allowing for easy access, reproducible, and collaborative engagement with the research findings. The repository is open and available under: \texttt{https://github.com/richardhoehn/mtsu.coms.qualifying-exam}\cite{Hoehn_Improving_Emotion_Detection_2023}
\clearpage

\addcontentsline{toc}{section}{References}
\printbibliography
\clearpage

\addcontentsline{toc}{section}{List of Figures}
\listoffigures
\clearpage

% Appendix Details
\appendix
\input{appendix/dataset} % Note: SubSection headers are handled within the *.tex file 
\input{appendix/code} % Note: SubSection headers are handled within the *.tex file

\end{document}