### 2023-08-14 paper/appendix/code.tex
Paper Updates | Savings States

\section{Appendix -- Application \& Code Details}
In this appendix we aim to supply details to the code and application blocks of this project. It should be noted that some of the code has been edited to fit on these pages and might not represent verbatim the code in the GIT repository.

All code was developed and test using a MacBook PRO running \texttt{Ventura 13.4.1}. Since PySpark was developed for cluster processing it may seem counter intuitive to build PySpark code for a Mac; however the intent is that some of this research can be further used on clusters in the future.

\subsection{PySpark Session Setup}
\label{appendix:py-spark-setup}
\begin{minted}[fontsize=\small]{python}
# Setup Spark Session
from pyspark.sql import SparkSession

# Get Imports Needed
from pyspark.sql.functions import col, udf, regexp_replace, lower
from pyspark.sql import functions as F

# Get Datatypes needed for DataFrame manipulation
from pyspark.sql.types import IntegerType, StringType, ArrayType

# Other Imports
import re  # Import the "re" module for regular expressions

# Setup Spark Session
sc = SparkSession \
        .builder \
        .master("local[*]") \
        .appName("Qualyfing-Exam") \
        .getOrCreate()

# Print Spark Version being run
print("Spark V: ", sc.version)
\end{minted}
\clearpage

\subsection{RESTful API Server}
\label{appendix:code-api-server}
The RESTful\footnote{RESTful API is an interface that two computer systems use to exchange information securely over the internet.} API Server will be a simple Python Flask server listening on port \texttt{8080} on the local machine.

Flask is a lightweight and simple to use web framework in Python, allowing developers to create web applications and APIs. Due to it's simplicity and components design, Flask allowed us to rapidly build and interactive web service fro demonstration purposes on this project.
\begin{minted}[fontsize=\small]{python}
# Setup Spark Session
from pyspark.sql import SparkSession

# Import API Modules 
from flask import Flask, request, jsonify
from flask_cors import CORS
import datetime

# Import Translator
from googletrans import Translator

# Impport Tokenizer and Pipeline Tools
from transformers import DistilBertTokenizer
from pyspark.ml.pipeline import PipelineModel

# Setup Spark Session
sc = SparkSession \
        .builder \
        .master("local[*]") \
        .appName("API-Demo") \
        .getOrCreate()

# Add Numerical Label for Emotions
emotion_key = {
    "boredom": 0,
    "love": 1,
    "relief": 2,
    "fun": 3,
    "hate": 4,
    "neutral": 5,
    "anger": 6,
    "happiness": 7,
    "surprise": 8,
    "sadness": 9,
    "worry": 10,
    "enthusiasm": 11,
    "empty": 12,
    "---": 13
}

# Initialize the AutoTokenizer
tokenizer = DistilBertTokenizer.from_pretrained("distilbert-base-multilingual-cased")

# Load Saved Models
model_en_rfc = PipelineModel.load("./models/model_en_rfc.model")
model_en_nb = PipelineModel.load("./models/model_en_nb.model")
model_en_lr = PipelineModel.load("./models/model_en_lr.model")
model_de_rfc = PipelineModel.load("./models/model_de_rfc.model")
model_de_nb = PipelineModel.load("./models/model_de_nb.model")
model_de_lr = PipelineModel.load("./models/model_de_lr.model")

app = Flask(__name__)
CORS(app)  # Use the CORS class to enable CORS for the entire app
translator = Translator()

# Setup Emotion Key Reverse
emotions = {v: k for k, v in emotion_key.items()}

@app.route('/predict', methods=['GET'])
def transform_data():
    start_time = datetime.datetime.now()
    sentence_en = request.args.get('sentence_en')
    sentence_de = request.args.get('sentence_de')

    print(f"English: {sentence_en}")
    print(f"German: {sentence_de}")

    if sentence_en:
        sentence_de = translator.translate(sentence_en, dest="de").text
    elif sentence_de:
        sentence_en = translator.translate(sentence_de, dest="en").text
    else:
        return jsonify({'error': 'No Sentence provided!'})

    print(f"English: {sentence_en}")
    print(f"German: {sentence_de}")

    words_en = tokenizer.tokenize(sentence_en)
    words_de = tokenizer.tokenize(sentence_de)

    # Create Englsih & German Dataframe
    df_en_api = sc.createDataFrame([(sentence_en, words_en)], ["sentence", "words"])
    df_de_api = sc.createDataFrame([(sentence_de, words_de)], ["sentence", "words"])

    predictions_en_rfc = model_en_rfc.transform(df_en_api)
    predictions_en_nb = model_en_nb.transform(df_en_api)
    predictions_en_lr = model_en_lr.transform(df_en_api)
    predictions_de_rfc = model_de_rfc.transform(df_de_api)
    predictions_de_nb = model_de_nb.transform(df_de_api)
    predictions_de_lr = model_de_lr.transform(df_de_api)

    # Get End Time
    end_time = datetime.datetime.now()
    elapsed_time = (end_time - start_time).total_seconds()

    #Build Response
    resp = {
        "sentence": {
            "english": sentence_en,
            "german": sentence_de,
        },
        "predictions": {
            "en_rfc": emotions.get(predictions_en_rfc.collect()[0]["prediction"]),
            "en_nb":  emotions.get(predictions_en_nb.collect()[0]["prediction"]),
            "en_lr":  emotions.get(predictions_en_lr.collect()[0]["prediction"]),
            "de_rfc": emotions.get(predictions_de_rfc.collect()[0]["prediction"]),
            "de_nb":  emotions.get(predictions_de_nb.collect()[0]["prediction"]),
            "de_lr":  emotions.get(predictions_de_lr.collect()[0]["prediction"]),
        },
        "metadata": {
            "spark": sc.version,
            "datetime": {
                "start": start_time,
                "end": end_time,
                "elapsedInSeconds": elapsed_time,
            }
        }
    }

    return jsonify(resp)

if __name__ == '__main__':
    app.run(host='127.0.0.1', port=8080)

\end{minted}
\clearpage